---
title: "Instacart Interactivity"
---

# Data description

We are going to use [Instacart](instacart_train_data.csv.zip) data for Homework 6. The dataset contains 1,384,617 observations of 131,209 unique users, where each row in the dataset is a product from an order. There is a single order per user in this dataset.

There are 15 variables in this dataset:

* `order_id`: order identifier
* `product_id`: product identifier
* `add_to_cart_order`: order in which each product was added to cart
* `reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
* `user_id`: customer identifier
* `eval_set`: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train”  eval_set)
* `order_number`: the order sequence number for this user (1=first, n=nth)
* `order_dow`: the day of the week on which the order was placed
* `order_hour_of_day`: the hour of the day on which the order was placed
* `days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
* `product_name`: name of the product
* `aisle_id`: aisle identifier
* `department_id`: department identifier
* `aisle`: the name of the aisle
* `department`: the name of the department

The code below loads the libraries and imports the data.

```{r eval=FALSE}
library(flexdashboard)
library(tidyverse)
library(janitor)
library(stringr)
library(forcats)
library(viridis)
library(plotly)
```

```{r eval=FALSE}
## Import and clean the data "Instacart".
instacart = read_csv("./instacart_train_data.csv.zip") %>%
  clean_names()
```


# Flexdashboard for Instacart

The flexdashboard using plotly for interactive graphics are [here](instacart_dashboard1.html).

## Plotly Barplot

First, we create a barplot to show how many items ordered for each department?

```{r eval=FALSE}
instacart %>% 
  count(department) %>% 
  rename(items = n) %>% 
  mutate(department = fct_reorder(department,items)) %>% 
  plot_ly(x = ~department, y = ~items, color = ~department, type = "bar")
```


## Plotly Boxplot

We are also interested that when people buy groceries online. We create a boxplot to show the distribution of the order hour of the day for each department.

```{r eval=FALSE}
instacart %>%
  mutate(department = fct_reorder(department, order_hour_of_day, IQR)) %>%
  select(department, order_hour_of_day) %>%
## The IQR of each department was in increasing order.
  plot_ly(y = ~order_hour_of_day, color = ~department, type = "box", colors = "Set2")
```

## Plotly scatterplot

We also use the code below to create a scattarplot to explore the association between number of orders and probability of reordering

```{r eval=FALSE}
## reorder
reordered = instacart %>% 
  filter(reordered == 1) %>% 
  count(product_id) %>% 
  rename(reordered = n)

instacart %>% 
  count(product_id, department) %>%
  rename(ordered = n) %>% 
  left_join(reordered) %>% 
  mutate(reordered = ifelse(is.na(reordered),0,reordered),
         reo_ratio = reordered/ordered,
         tlabel = str_c("Product_ID: ", product_id, product_id, '\nDepartment: ', department)) %>%
  plot_ly(x = ~ordered,y = ~reo_ratio, type = "scatter", mode = "markers", alpha = 0.5, text = ~tlabel, color = I("black"))
```


# Shiny-dashboard for Instacart

The Shiny dashboard are hosted [here](https://ashleyyyw.shinyapps.io/instacart_dashboard2/).

Based on the plots in the flexdashboard, we add a `selectInput` to choose department and a `sliderInput` widget to select the range of how many items are ordered.

```{r eval=FALSE}
dpts = instacart %>% distinct(department) %>% pull()

# selectInput widget
selectInput("dpt_choice", label = h3("Select department"), 
            choices = dpts, selected = "dairy eggs")

item_n = instacart %>% count(aisle)
instacart = left_join(instacart, item_n, "aisle") %>% rename(item_n = n)
max_item = max(item_n$n)
min_item = min(item_n$n)

# sliderInput widget
sliderInput("item_count", label = h3("Range of item counts"), min = min_item, 
        max = max_item, value = c(10000, 50000))
```

To make the plots interactive, we replace the renderText with the code below in our flexdashboard.

The scatterplot

```{r eval=FALSE}
reordered = instacart %>% 
  filter(reordered == 1) %>% 
  count(product_id) %>% 
  rename(reordered = n)

renderPlotly({
  instacart %>% 
  filter(department == input$dpt_choice,
         item_n %in% input$item_count[1]:input$item_count[2]) %>% 
  count(product_id, department) %>%
  rename(ordered = n) %>% 
  left_join(reordered) %>% 
  mutate(reordered = ifelse(is.na(reordered),0,reordered),
         reo_ratio = reordered/ordered,
         tlabel = str_c("Product_ID: ", product_id, product_id, '\nDepartment: ', department)) %>%
  plot_ly(x = ~ordered,y = ~reo_ratio, type = "scatter", mode = "markers", alpha =0.5, text = ~tlabel, color = I("black"))
})
```

The barplot

```{r eval=FALSE}
renderPlotly({
  instacart %>% 
  filter(department == input$dpt_choice,
         item_n %in% input$item_count[1]:input$item_count[2]) %>%
  mutate(aisle = fct_reorder(aisle, item_n)) %>% 
  plot_ly(x = ~aisle, y = ~item_n, color = ~aisle, type = "bar")
})

```

The boxplot

```{r eval=FALSE}
renderPlotly({
  instacart %>%
  filter(department == input$dpt_choice,
         item_n %in% input$item_count[1]:input$item_count[2]) %>% 
  mutate(aisle = fct_reorder(aisle, order_hour_of_day, IQR)) %>%
  select(department, aisle, order_hour_of_day) %>%
  plot_ly(y = ~order_hour_of_day, color = ~aisle, type = "box", colors = "Set2")
})
```

You can access to the GitHub repository [here](https://github.com/ashleyyyw/interactivity_practice).

<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
h1 { /* Header 1 */
  font-size: 28px;
}
h2 { /* Header 2 */
  font-size: 18px;
}
</style>